Exp 2 of 4

### 2016-08-12 experiments/e2/writeup/e2-b-knitr.Rnw
more consistency, better plots, using discriminability instead of item ratio

\documentclass[a4paper]{article}
\usepackage{geometry}
\usepackage{placeins}
\begin{document}

<<loadKnitr, echo=FALSE, message=FALSE>>=
library(knitr)
opts_chunk$set(cache=FALSE, autodep = TRUE)
@

<<defineFonts>>=
# .define.fonts <- function () {
#     quartzFonts(avenir =    c('Avenir Book', 'Avenir Black', 'Avenir Book Oblique', 'Avenir Black Oblique'),
#                 helvetica = c('Helvetica New Light', 'Helvetica New Bold', 'Helvetica New Light Italic', 'Helvetica New Bold Italic'),
#                 cmodern = c('CMU Serif', 'CMU Serif Extra','CMU Classical Serif','CMU Serif Extra'))
# }
# .define.fonts()
@

<<loadFonts>>=
library(extrafont)
loadfonts()
@

<<loadLibs, echo=FALSE, message=FALSE>>=
library(beepr)                     #
library(car)                       #
library(ggplot2)                   #
library(grid)                      #
library(gridExtra)                 #
library(languageR)                 #
library(lattice)                   #
library(latticeExtra)              # for doubleYScale
library(lme4)                      #
library(lmerTest)                  #
library(LMERConvenienceFunctions)  # for perSubjectTrim
library(MASS)                      #
library(plyr)                      #
library(reshape2)                  #
library(rms)                       # for varclus
library(Rmisc)                     # for summarySEwithin and friends
library(xtable)                    #
@

\section{Get the data}

<<getTheData, echo=FALSE>>=
# if the file data.Rda already exists then load it, else do data wrangling
if( file.exists('data_processed.Rda') ) {
	load('data_processed.Rda')
	} else {
	gatherData = function(number_of_valid_subjects) {
		data_dir <- '../experimentCode/output/'
		column.headers.df <-  head( read.table(
			paste(data_dir,'subject01.data',sep=''),
			header=TRUE),0)
		gathered.data <- column.headers.df
		for (subject in 1:number_of_valid_subjects) {
			current.filename <- paste(data_dir, 'subject', sprintf("%02d",subject), '.data', sep='')
			current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = FALSE)
			gathered.data <- rbind(gathered.data, current.data)
		return(gathered.data)
	} # end of gatherData function
	classifyResponse = function(dat) {

@@

						 'NOANSWER' = {dat[row, 'response_num'] <- NA}
			)

@@

@@

	# declare local variables
	number_of_valid_subjects <- 30 # = 30
	number_of_rows <- 7680 # 7680
	number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # = 256
	dat <- gatherData(number_of_valid_subjects) # = 30
	dat <- classifyResponse(dat) # classify the response as expected, near, or far

@@

	dat$id <- factor(paste(paste(dat$Subject),
												 paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))

@@

	dat$c_Itm <- ifelse(dat$Item==1, -.75,
											ifelse(dat$Item==2, -.25,
														 ifelse(dat$Item==3, .25, .75)))

@@

	# add ratios for Item
	item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54)
	# 0.2500000 0.4705882 0.5909091 0.6666667
	item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
	# -1.3441995 -0.1316642  0.5297187  0.9461450
	item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))
	# 0.5125000 0.6876471 0.7691558 0.8166667
	item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))))
	# -1.37582241 -0.06614191  0.54334858 0.89861574
	dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
	dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
	dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
	dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667
	dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
	dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
	dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187
	dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450
	dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000
	dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471
	dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
	dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667
	dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241
	dat[dat$Item == "16:25:34", "item_mean_ratio_scaled"] <-   -0.06614191
	dat[dat$Item == "26:35:44", "item_mean_ratio_scaled"] <-    0.54334858
	dat[dat$Item == "36:45:54", "item_mean_ratio_scaled"] <-    0.89861574

@@

	dat$RT_RecSqd    <- 1/dat$RT^2
	dat$RT_Rec       <- 1/dat$RT
	dat$RT_RecSqt    <- 1/sqrt(dat$RT)

@@

	dat$RT_sqt       <- sqrt(dat$RT)

@@

	dat$RT_sqd       <- dat$RT^2

@@

	design_info <- unique(subset(dat, select=c(Item, Condition, Vagueness, Number, Quantity, Order,
																						 Left, Mid, Right, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side,Instruction)))

@@

	capture.output(print.data.frame(design_info, row.names=F, print.gap=3, quote=F, right=F),
								 file="design_info.table")

@@

	dat <- subset(dat,
								select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number,
													 Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,

@@

													 RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
													 Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
													 Instruction, nchar_instr
								))
	# This data set (dat) contains *all* trials 7680 including impossible trials and is
	# mainly for graphs comparing different removals

@@

	# remove impossible trials from dat in new data dd #
	# Throw out RT = 1 and RT = 59998, and RTprev = 1 and RTprev = 59998
	# i.e., throw out sticky fingers and timeouts,
	# and the trials that followed sticky fingers and timeouts
	# since they were likely affected by unusual previous trials.
	# lose impossible trials
	dd <- dat
	dd$RT[dd$RT == 1 ] <- NA
	dd$RT[dd$RT == 59998 ] <- NA
	dd <- dd[complete.cases(dd),]
	row.names(dd) <- NULL
	# add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
	dd$RTprev <- NA
	for (s in levels(dd$Subject)) {
		nrows = nrow(dd[dd$Subject==s,])
		for (i in 1:nrows) {
			if (i==1) {
				dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
			else
				dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
	# add transformations of previous RT
	dd$RTprev_RecSqd    <- 1/dd$RTprev^2
	dd$RTprev_Rec       <- 1/dd$RTprev
	dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev)
	dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
	dd$RTprev_log       <- log(dd$RTprev)
	dd$RTprev_sqt       <- sqrt(dd$RTprev)
	dd$RTprev_raw       <- dd$RTprev
	dd$RTprev_sqd       <- dd$RTprev^2
	# number of higher-level cells in the design (abstracting over Order) =
	# Item(4) * Number(2) * Vagueness(2) * Quantity(2) = 32
	# number of measurements in eaach cell = 8
	# 332 * 8 = 256
	# 256 * 30 = 7680
	dd$measurement=0
	dd$cell=0
	for (s in levels(dd$Subject)) {
		cellcount=0
		for (i in levels(dd$Item)) {
			for (n in levels(dd$Number)) {
				for (v in levels(dd$Vagueness)) {
					for (q in levels(dd$Quantity)) {
						measurementcount=0
						cellcount=cellcount+1
						for (t in dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q, "Trial"]) {
							dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "cell"] <- cellcount
							measurementcount=measurementcount+1
							dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "measurement"] = measurementcount
	# put dd in better column order
	dd <- subset(dd,
							 select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number,
							 					 Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
							 					 RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
							 					 RTprev, RTprev_RecSqd, RTprev_Rec, RTprev_RecSqt, RTprev_log, RTprev_sqt, RTprev_raw, RTprev_sqd,
							 					 Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
							 					 Instruction, nchar_instr,
							 					 cell, measurement
							 ))
	save(dd, file="data_processed.Rda")
} # end of the else clause, which is run if the file data.Rda does not exist.

@@

\section{Plots}

<<prepPlotData, echo=FALSE>>=
# first create a data frame specially for transformations
# this won't be used for analysis, just for this plot
load('data_raw.Rda') # dat
dat_transforms <- dat
# dfSubset says whether any data points were removed
dat_transforms$dfSubset <- "original"
# RTtype says whether the row contains data for an observed RT or 
# a sample from a normal distribution with the same mean and sd
dat_transforms$RTtype <- "observed"
# create extra rows for dfs with removed (outlier) data points
dat3 <- subset(dat_transforms, RT>1 & RT<59999)              # nrow 7677
dat3$dfSubset <- "RT>1&RT<59999"
dat4 <- subset(dat_transforms, RT>1 & RT<59999 & RT<30001)   # nrow 7658
dat4$dfSubset <- "RT>1&RT<30001"
# bring the dfs back into main df
dat_transforms <- rbind(dat3, dat4)
dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("RT>1&RT<59999", "RT>1&RT<30001"))
# create extra rows for normally distributed transformed variables
dato = dat_transforms
datn = dat_transforms; datn$RTtype <- "normal"
dat_transforms <- rbind(dato,datn)
dat_transforms$RTtype <- factor(dat_transforms$RTtype, levels=c("observed","normal"))
# for safety
dat_transforms[dat_transforms$RTtype=="normal", c("RT_log", "RT_raw")] <- NA
# Create the normally distributed transformed variables, replacing the NAs declared above
for (k in levels(dat_transforms$dfSubset)){
    dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_log"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_log"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_log"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_log"]))
    dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_raw"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_raw"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_raw"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_raw"]))
  }# end of create normally distributed equivalents for the data transforms.
temp <- reshape2::melt(dat_transforms, 
             id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Number", "Quantity", "Order"),
             measure.vars=c("RT_log", "RT_raw"),
             variable.name="transformation",
             value.name="time")
@

<<plotDists, fig.width=5, fig.height=5, fig.cap='Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions', fig.pos='hbtp', fig.align='right', message=FALSE, warning=FALSE, echo=FALSE>>=
ggplot(temp, aes(time, colour=RTtype)) + 
  geom_freqpoly() + 
  facet_wrap(transformation~dfSubset, scales='free') + 
  scale_color_manual(values=c("blue","red")) + 
  theme_bw() + 
  theme(aspect.ratio=1, text=element_text(family='CMU Serif', size=12, face='italic'))
@

<<exampleOfFonts, fig.width=3, fig.height=3, fig.cap='font example', echo=FALSE>>=
p <- ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point() +
ggtitle("Fuel Efficiency of 32 Cars") +
xlab("Weight (x1000 lb)") + ylab("Miles per Gallon") +
theme_bw() +
theme(text=element_text(family="CMU Serif", size=14, face='plain'))
p
ggsave("ggplot_CMU_Serif.pdf", p, width=3.5, height=3.5)
@

<<boxPlots, fig.width=7, fig.height=3, fig.cap='Show how transformations of RT affect distribution of times, and how they affect which times are outliers.', echo=FALSE>>=
load('data_processed.Rda')
# allow reference to transformation
subdata = reshape2::melt(dd,
                         measure.vars=c("RT_log", "RT_raw"),
                         variable.name="transformation",
                         value.name="value") 
# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ transformation, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=5), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="orange","Subjects"="blue")) +
  xlab("") + ylab("")
  @
  <<fastSlowRT, fig.width=7, fig.height=3, fig.cap='Show how mean times for individual subjects and items vary with respect to the grand mean Log RT.', message=FALSE, echo=FALSE>>=
# Show how mean times for individual subjects and items vary with respect to the grand mean RT.
subs=ddply(dd, .(level=Subject), 
summarise, effect="subject", emean=mean(dd$RT_log), lmean=mean(RT_log),
ldiff=mean(RT_log)-mean(dd$RT_log))
itms=ddply(dd, .(level=Item), 
summarise, effect="item", emean=mean(dd$RT_log), lmean=mean(RT_log), 
ldiff=mean(RT_log)-mean(dd$RT_log))
si=rbind(subs,itms)
si$effect <- factor(si$effect, levels=c("subject", "item"))
my1 <- xyplot(data=si, lmean~level | effect, 
aspect=1,
layout=c(2,1),
ylim = extendrange(si$lmean, f = 0.10),
par.settings=list(trellis.par.set(superpose.symbol = list(col='black'), 
superpose.line=list(col='black'))),
xlab=NULL,
ylab=NULL,
main='How mean times for individual subjects and items\nvary with respect to the grand mean Log RT',
mean=unique(si$emean),
scales=list(x=list(relation="free", at=NULL, 
labels=list("subject number", "item identifier"), cex=1 ),
y=list(at=    c(0,7,unique(si$emean),8,9),
labels=c("0","7","Mean Log RT","8","9"))),
panel=function(x,y,type,subscripts,...){
panel.abline(h=unique(si$emean), col="lightgrey",lty=1,...)
panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, subscripts=subscripts, code=3, length=0, ...)
panel.xyplot(x,y,type='p', pch=21, cex=.5, col='black', fill='black', subscripts=subscripts,...)
}) # end of my1
my2 <- xyplot(data=si, ldiff~level | effect, 
aspect=1, type='n',
ylab="Deviation from\nmean Log RT",
ylim=extendrange(si$ldiff,f=0.10)
) # creates a second y-axis scale
print(doubleYScale(my1, my2, 
add.axis=TRUE, 
add.ylab2=TRUE, 
use.style=FALSE))
@

<<plotMainsa, fig.width=6, fig.height=6, fig.cap='Plot main effects on several transformations', message=FALSE, echo=FALSE, warning=FALSE>>=
# plot main effects on observed data
# allow reference to transformation
temp <- reshape2::melt(dd,
id.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log",  "RT_raw"),
variable.name="transformation",
value.name="score")
# allow reference to effect
temp2 <- reshape2::melt(temp,
id.vars=c("transformation", "score"),
measure.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
variable.name="effect",
value.name="level")
# do the plot
ggplot(temp2, aes(y=score, x=level, group=1, col=effect)) +
facet_wrap(~transformation + effect, scales="free", shrink=TRUE, ncol=5) +
stat_summary(fun.data=mean_cl_normal, geom="pointrange") +
stat_summary(fun.y=mean, geom="line") +
theme(aspect.ratio=1, axis.text.y=element_blank())
@

<<mainB, fig.width=6, fig.height=3, fig.cap='Just the main effects', message=FALSE, echo=FALSE>>=
# prep main effects rts data frame
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number")));names(nrts)[3]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Order")));names(orts)[3]<-"Levels"
irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Item")));names(irts)[3]<-"Levels"
rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci
# plot main effects
ggplot(rts, aes(y=RT_log, x=Levels, group=1, ymin=mins, ymax=maxs)) +
geom_line() +
geom_errorbar(width=.1) +
geom_point(cex=3) +
facet_grid(~Effect, scale="free_x") +
theme_bw() +
theme(aspect.ratio=1)
@

<<mainD, fig.width=6, fig.height=3.5, fig.cap='Main effects over item ratios', message=FALSE, echo=FALSE>>=
nrts <- cbind(Effect="Number",
xlev=factor(c(1,2)), 
myshape='a',
summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Number")))
names(nrts)[5]<-"Levels"

orts <- cbind(Effect="Order",
xlev=factor(c(1,2)), 
myshape='b',
summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Order")))
names(orts)[5]<-"Levels"

>>>>

\documentclass[a4paper,12pt]{article}
\usepackage[margin=2cm]{geometry}
\usepackage{placeins}

\title{Exp 2 of 4}
\begin{document}

<<setup, include=FALSE, cache=FALSE>>=
library(knitr)
options(formatR.arrow=TRUE, 
        width=100)
opts_chunk$set(fig.path='figure/graphics-', 
               cache.path='cache/graphics-', 
               fig.align='center', 
               dev='pdf', 
               fig.width=5, 
               fig.height=5, 
               fig.show='hold', 
               cache=FALSE, 
               par=TRUE,
               size='scriptsize',
               tidy=TRUE,
               comment = NA,
               tidy.opts = list(blank=FALSE, 
                                width.cutoff=100))
knit_hooks$set(crop=hook_pdfcrop)
@

\maketitle
\tableofcontents
\clearpage

<<loadFonts, message=FALSE, echo=FALSE>>=
library(fontcm)
library(extrafont)
#font_import()
#loadfonts()
@

<<loadLibs, message=FALSE, echo=FALSE>>=
library(beepr)                     #
library(car)                       #
library(ggplot2)                   #
library(grid)                      #
library(gridExtra)                 #
library(languageR)                 #
library(lattice)                   #
library(latticeExtra)              # for doubleYScale
library(lme4)                      #
library(lmerTest)                  #
library(LMERConvenienceFunctions)  # for perSubjectTrim
library(MASS)                      #
library(plyr)                      #
library(reshape2)                  #
library(rms)                       # for varclus
library(Rmisc)                     # for summarySEwithin and friends
library(xtable)                    #
@

<<gatherDataFunction, echo=FALSE>>=
gatherData = function(number_of_valid_subjects) {
  data_dir <- '../experimentCode/output/'
  column.headers.df <-  head( read.table(
    paste(data_dir,'subject01.data',sep=''),
    header=TRUE),0)
  gathered.data <- column.headers.df
  for (subject in 1:number_of_valid_subjects) {
    current.filename <- paste(data_dir, 'subject', sprintf("%02d",subject), '.data', sep='')
    current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = FALSE)
    gathered.data <- rbind(gathered.data, current.data)
  }
  message('Returning gathered data')
  return(gathered.data)
} # end of gatherData function
@

<<classifyResponseFuntion, echo=FALSE>>=
classifyResponses = function(dat) {

@@

           'NOANSWER' = {dat[row, 'response_num'] <- NA})}

@@

  message('Returning classified data')

@@

<<processDataFunction, echo=FALSE>>=
processData = function(dat) {

@@

  dat$id <- factor(paste(paste(dat$Subject), paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))

@@

  dat$c_Itm <- ifelse(dat$Item==1, -.75, ifelse(dat$Item==2, -.25, ifelse(dat$Item==3, .25, .75)))

@@

@@

@@

@@

@@

  design_info <- unique(subset(dat, select=c(Item, Condition, Vagueness, Number, Quantity, Order, Left, Mid, Right, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, Instruction)))

@@

  capture.output(print.data.frame(design_info, row.names=F, print.gap=3, quote=F, right=F), file="design_info-table.txt")
  # Add discriminability metric with reference to item
  discriminability_range = c(0.7500000, 0.5294118, 0.4090909, 0.3333333)
  discriminability_range_scaled = c(1.3441995, 0.1316642, -0.5297187, -0.9461450)  
  discriminability = c(0.4875000, 0.3123529, 0.2308442, 0.1833333)
  discriminability_scaled = c(1.37582241, 0.06614191, -0.54334858, -0.89861574)
  dat[dat$Item == "06:15:24", "discriminability"] <- 0.4875000
  dat[dat$Item == "16:25:34", "discriminability"] <- 0.3123529
  dat[dat$Item == "26:35:44", "discriminability"] <- 0.2308442
  dat[dat$Item == "36:45:54", "discriminability"] <- 0.1833333

@@

  dat <- subset(dat, select = c(id, Subject, Trial, Condition, Order, 
                           Quantity, Vagueness, Number, Item, discriminability, 

@@

                           RT, RT_log, RT_raw, Exp_Num, Bline_Num, 
                           Extr_Num, Exp_side, Bline_side, Extr_side, 
                           response_num, response_side, response_category, 
                           Left, Mid, Right, Instruction, nchar_instr))
  # This data set (dat) contains *all* trials 7680 including impossible trials and is mainly for graphs comparing different removals

@@

  message('Returning processed data')
} # end of function processData

@@

<<postProcessDataFunction, echo=FALSE>>=
postProcessData = function(dat) {
  # dd removes impossible trials from dat
  # Throw out RT = 1 and RT = 59998, and RTprev = 1 and RTprev = 59998 i.e., throw out sticky fingers and timeouts, and the trials that followed sticky fingers and timeouts since they were likely affected by unusual previous trials. Also lose impossible trials
  dd <- dat
  dd$RT[dd$RT == 1 ] <- NA
  dd$RT[dd$RT == 59998 ] <- NA
  dd <- dd[complete.cases(dd),]
  row.names(dd) <- NULL
  # add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
  dd$RTprev <- NA
  for (s in levels(dd$Subject)) {
    nrows = nrow(dd[dd$Subject==s,])
    for (i in 1:nrows) {
      if (i==1) {dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]}
      else
        dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1] }}
  # add transformations of previous RT
  dd$RTprev_log       <- log(dd$RTprev)
  dd$RTprev_raw       <- dd$RTprev
  # put dd in better column order
  dd <- subset(dd, select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number, Item, discriminability, c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty, RT, RT_log, RT_raw, RTprev, RTprev_log, RTprev_raw, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right, Instruction, nchar_instr ))
  save(dd, file="data_processed.Rda")
  message('Returning post processed data')
  return(dd)
} # end of function postProcessResponses
@

\clearpage
\section{Get the data}

<<getTheData>>=
# if the file data_processed.Rda already exists then load it, else do data wrangling
if( file.exists('data_processed.Rda') ) {
  load('data_processed.Rda')
} else {
  # declare local variables
  number_of_valid_subjects <- 30 # = 30
  number_of_rows <- 7680 # 7680
  number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # = 256
  # call functions
  dat <- gatherData(number_of_valid_subjects) # = 30
  dat <- classifyResponses(dat) # classify the response as expected, near, or far
  dat <- processData(dat) # remove impossible trials and re-do previous rt measures
  dd  <- postProcessData(dat)
} # end else do data wrangling
@

<<showDataNames>>=
names(dd)
@

<<showDataHead>>=
head(dd)
@

<<showDataSummary>>=
summary(dd)
@

<<showDataStructure>>=
str(dd)
@

\clearpage
\section{Plots}
\clearpage
\subsection{Discriminability}

<<showCompression, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='Ratios for different numbers of dots in the arrays: smaller values are more discriminable. Blue is for the ratio between the smallest number in the array and the largest number in an array. Red is for the mean of two ratios, one for the smallest number to the middle number, the other for the middle number to the largest number in the array', echo=FALSE>>=
#  Show compression of ratios of numbers of dots in the items
discriminability = sort(unique(as.numeric(dd$discriminability)))
plot(y=c(1.5,1.5,1.5,1.5), 
     x=discriminability, 
     ylab='',
     xlab='',
     yaxt='n',
     type='n', 
     xaxt='n',
     axes=F)
lines(y=c(1.5,1.5,1.5,1.5), x=discriminability, type='o', pch=19)
text(y=c(1.15,1.15,1.15,1.15), x=discriminability, labels=round(discriminability,2),  cex=.7)
mtext('mean (ratio min/med,\nratio med/max)', side = 2, line = 1)
mtext('Discriminability index', side=1, line=3)
mtext('Discriminability for each item', side=3, line=1, cex=1.5)
axis(1)
box(col='grey')
#text(y=1.00, x=0.07, labels="mean\n(ratio min/med,\nratio med/max)", col='red')
# axis(1, at=seq(-.1, 1,0.1)[4:11])
# axis(1)
@

\clearpage

<<prepPlotData, echo=FALSE>>=
# first create a data frame specially for transformations
# this won't be used for analysis, just for this plot
load('data_raw.Rda') # dat
dat_transforms <- dat
# dfSubset says whether any data points were removed
dat_transforms$dfSubset <- "original"
# RTtype says whether the row contains data for an observed RT or 
# a sample from a normal distribution with the same mean and sd
dat_transforms$RTtype <- "observed"
# create extra rows for dfs with removed (outlier) data points
dat3 <- subset(dat_transforms, RT>1 & RT<59999)              # nrow 7677
dat3$dfSubset <- "RT>1&RT<59999"
dat4 <- subset(dat_transforms, RT>1 & RT<59999 & RT<30001)   # nrow 7658
dat4$dfSubset <- "RT>1&RT<30001"
# bring the dfs back into main df
dat_transforms <- rbind(dat3, dat4)
dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("RT>1&RT<59999", "RT>1&RT<30001"))
# create extra rows for normally distributed transformed variables
dato = dat_transforms
datn = dat_transforms; datn$RTtype <- "normal"
dat_transforms <- rbind(dato,datn)
dat_transforms$RTtype <- factor(dat_transforms$RTtype, levels=c("observed","normal"))
# for safety
dat_transforms[dat_transforms$RTtype=="normal", c("RT_log", "RT_raw")] <- NA
# Create the normally distributed transformed variables, replacing the NAs declared above
for (k in levels(dat_transforms$dfSubset)){
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_log"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_log"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_log"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_log"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_raw"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_raw"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_raw"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_raw"]))
}# end of create normally distributed equivalents for the data transforms.
@

<<plotDistFreqPolys, fig.width=6, fig.height=4, fig.cap='Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions', fig.pos='hbtp', fig.align='center', message=FALSE, warning=FALSE, echo=FALSE>>=
temp <- reshape2::melt(dat_transforms, 
                       id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Number", "Quantity", "Order"),
                       measure.vars=c("RT_log", "RT_raw"),
                       variable.name="transformation",
                       value.name="time")
ggplot(temp, aes(time, colour=RTtype)) + 
  geom_freqpoly() + 
  facet_wrap(transformation~dfSubset, scales='free', ncol=4, nrow=1) + 
  scale_color_manual('RT Type: ', values=c("blue","red")) + 
  theme_bw() + 
  theme(aspect.ratio=1, 
        text=element_text(family='CMU Serif', size=10, face='plain'),
        axis.title.x=element_blank(),
        axis.text.x=element_text(angle=25),
        panel.grid=element_blank(),
        legend.position='top',
        legend.key=element_rect(color = 'white'))
@

\clearpage

<<plotBoxPlots, fig.width=6, fig.height=4, fig.cap='Show how transformations of RT affect distribution of times, and how they affect which times are outliers.', fig.pos='hbtp', echo=FALSE>>=
load('data_processed.Rda')
# allow reference to transformation
subdata = reshape2::melt(dd,
                         measure.vars=c("RT_log", "RT_raw"),
                         variable.name="transformation",
                         value.name="value") 
# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ transformation, scales="free", ncol=2, nrow=1) +
  geom_boxplot(aes(y=value, x=Item, col="Items"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
  theme_bw() +
  theme(aspect.ratio=1, 
        text=element_text(family='CMU Serif', size=10, face='plain'),
        panel.grid=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.x=element_blank(),
        legend.position='top',
        legend.key=element_rect(color = 'white')) + 
  scale_colour_manual("",values=c("Items"="orange","Subjects"="blue")) +
  xlab("") + ylab("")
@

\clearpage
